---
title: "STE:PHEN Skills"
author: "Alan Nielsen"
date: "September 19, 2017"
output: html_document
---

This R markdown document generates some graphs showing the various distributions of self-assessed skills of attendees

First we load a library and read in the data:

```{r Preparation, message = FALSE, warning = FALSE}

library(tidyverse)
library(reshape2)

skills <- read.csv("F:/Google Drive/GitHub Repos/Crossmodality-Toolkit/STEPHEN/StephenSkills.csv")

```

Then we need to put the data frame into long format for use by ggplot

```{r Melting, message = FALSE}
skillstemp <- subset(skills, select = -c(ID))
moltenskills <- melt(skillstemp, variable.name = "Skillset", id.vars = "Group")
```

Then we can make some pretty histograms

```{r Histograms, message = FALSE, warning = FALSE}

ggplot(data=moltenskills, aes(moltenskills$value, fill = Skillset, group= Skillset)) + 
  geom_histogram(binwidth = 1) +
  ggtitle("Self-Reported Skill Levels of September Tutorial Participants") +
  labs(x="Rating", y="Count") +
  theme(axis.title.y = element_text(size=12,  color="#666666")) +
  theme(axis.text = element_text(size=8)) +
  theme(plot.title = element_text(size=16, face="bold", hjust=0, color="#666666")) +
  facet_wrap(~ Skillset, nrow = 2) +
  theme(legend.position="none") +
  theme(strip.text.x = element_text(size = 7, colour = "black"))

```








That shows us the distribution of skills- but how about what we actually end up with per group

```{r Barplots, warning = FALSE, message = FALSE}
moltenskills$Skillset <- factor(moltenskills$Skillset, 
                                  level = c("Bayes.and.Agent.Based.Modelling", "Advanced.Statistics.in.R", "LMER.in.R", "Online.Experiments",
                                            "JavaScript", "Data.Handling", "Experiment.Design", "Sound.Symbolism"),
                                labels = c("Bayes and Modelling", "Advanced Statistics in R", "LMER in R", "Online Experiments",
                                            "JavaScript", "Data Handling", "Experiment Design", "Sound Symbolism"))


ggplot(data=moltenskills, aes(x= Skillset, y = value, fill = Skillset, group= Skillset)) + 
  geom_bar(stat= "summary") +
  ggtitle("Mean Skill Level by Group") +
  labs(x="Skillset", y="Mean Skill Level") +
  theme(axis.title.y = element_text(size=12,  color="#666666")) +
  theme(axis.text = element_text(size=8)) +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1))+
  theme(plot.title = element_text(size=16, face="bold", hjust=0, color="#666666")) +
  theme(legend.position="none") +
  theme(strip.text.x = element_text(size = 12, colour = "black")) +
  facet_wrap(~ Group, nrow = 3) +
  coord_flip()

```
